---
title: "p8105_hw2_zg2486"
output: github_document
date: "2024-09-27"
---

# Problem 1

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

1. Read and clean the data, select and convert variables.
```{r}
subway_df = 
  read_csv(
    "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
    na = c("NA", "", ".")) %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(
    entry = case_match(
      entry,
     "YES" ~ TRUE,
     "NO" ~ FALSE)
    ) 
```

Variables included in this dataset are: line, station name, station latitude, station longitude, served routes (route 1 - 11), entry, vending, entrance type, and ADA compliance. I used `janitor::clean_names` to clean up column names of data after importing data file with specifying missing values, then retained 19 required variables by `select` function, lastly, converted entry variable from character to logical variable by `mutate(case_match)` function. There are 1,868 rows and 19 columns.These data are not tidy yet.









