---
title: "p8105_hw2_zg2486"
output: github_document
name: Ziqi Guo
date: "2024-09-27"
---

# Problem 1

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

1. Read and clean the data, select and convert variables.
```{r}
subway_df = 
  read_csv(
    "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
    na = c("NA", "", "."), show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(
    entry = case_match(
      entry,
     "YES" ~ TRUE,
     "NO" ~ FALSE)
    ) 
```

Variables included in this dataset are: `line`, `station name`, `station latitude`, `station longitude`, `served routes (route 1 - 11)`, `entry`, `vending`, `entrance type`, and `ADA compliance`. I used `janitor::clean_names` to clean up column names of data after importing data file with specifying missing values, then retained `19` required variables by `select` function, lastly, converted entry variable from character to logical variable by `mutate(case_match)` function. There are `1,868` rows and `19` columns.These data are not tidy yet.

2. Answer 3 questions.
```{r eval=FALSE}
distinct_station = group_by(subway_df, station_name) %>% 
  distinct(line) %>% 
  nrow()


ada_station = filter(subway_df, ada == TRUE) %>% 
  group_by(station_name) %>% 
  distinct(line) %>% 
  nrow()

no_vending = filter(subway_df, vending == "NO") %>% 
  group_by(station_name) %>% 
  distinct(line) %>% 
  nrow()

entrance_allow = filter(subway_df, vending == "NO") %>% 
  filter(entry == TRUE) %>% 
  group_by(station_name) %>% 
  distinct(line) %>% 
  nrow()

proportion_entry = entrance_allow / no_vending
```
There are `465` distinct stations. and `84` stations are ADA compliant.`43.43%` of station entrances / exits without vending allow entrance.

3. Reformat data.
```{r}
str(subway_df)

reformat_df = subway_df %>%
  mutate(route8 = as.character(route8)) %>% 
  mutate(route9 = as.character(route9)) %>% 
  mutate(route10 = as.character(route10)) %>% 
  mutate(route11 = as.character(route11)) %>% 
  pivot_longer(
    route1:route9,
    names_to = "route_number",
    values_to = "route_name")

a_train = filter(reformat_df, route_name == "A") %>% 
  group_by(station_name) %>% 
  distinct(line) %>% 
  nrow()

ada_a_train = filter(reformat_df, ada == TRUE) %>% 
  filter(route_name == "A") %>% 
  group_by(station_name) %>% 
  distinct(line) %>% 
  nrow()
```
There are `r a_train` distinct stations serve the A train. There are `r ada_a_train` ADA compliant of the stations that serve the A train. 



# Problem 2

```{r message = FALSE}
mr_df = 
  read_excel("./202409 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel", na = c("NA", "", "."),
  ) %>% 
  select(-starts_with("...")) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(sports_balls = as.integer(sports_balls)) %>% 
  mutate(name = "Mr") 

professor_df = 
  read_excel("./202409 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel", na = c("NA", "", ".")
  ) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(name = "Professor")

gwynnda_df = 
  read_excel("./202409 Trash Wheel Collection Data.xlsx",
             sheet = "Gwynnda Trash Wheel", na = c("NA", "", ".")
  ) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(name = "Gwynnda")

wheel_tidy = 
  bind_rows(mr_df, professor_df, gwynnda_df) %>% 
  janitor::clean_names() %>% 
  relocate(name) 

total_observation = wheel_tidy %>% 
  nrow()

total_weight = 
  wheel_tidy %>% 
  filter(name == "Professor") %>% 
  summarise(total_weight = sum(weight_tons, na.rm = TRUE)) %>% 
  pull(total_weight)

total_number = 
  wheel_tidy %>% 
  filter(name == "Gwynnda", month == "June", year == 2022) %>% 
  summarise(total_number = sum(cigarette_butts)) %>% 
  pull(total_number)

```
There are total `r total_observation` observations in the combined dataset, the key variables include `date`, `weight_tons`, `volume_cubic_yards`, `plastic_bottles`, and `wrappers`, ect. The total weight of trash collected by Professor Trash Wheel is `r total_weight` tons. There total number of cigarette butts collected by Gwynda in June of 2022 are `r total_number`.



# Problem 3

1.Import and clean the data.
```{r}
bakers_df = 
  read_csv("./gbb_datasets/bakers.csv", 
           na = c("NA", "", "."), show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  mutate(baker = word(baker_name, 1)) %>% 
  select(-baker_name) 
  
bakes_df = 
  read_csv("./gbb_datasets/bakes.csv",
           na = c("NA", "", "."), show_col_types = FALSE) %>% 
  janitor::clean_names() 

results_df = 
  read_csv("./gbb_datasets/results.csv",
           na = c("NA", "", "."),
           skip = 2, show_col_types = FALSE) %>% 
  janitor::clean_names() 
```
I first use `clean_names` function to clean three data frames, and when we check the variable names of three data frames, we see there is a variable called `baker_name` that includes the full name of the contestants in baker_df instead of `baker`that only includes first name in other two data frame, which will hinder the merging of three data frames, so we will use `mutate` function to convert full name to first name and use `select` remove the full name column. 

2. Anti_join and merge datasets.
```{r}
str(bakers_df)
str(bakes_df)
str(results_df)

bakers_results_df = 
  anti_join(results_df, bakers_df, by = c("series", "baker" ))

results_df1 = results_df %>% 
  mutate(
    baker = ifelse(baker == "Joanne", "Jo", baker)) %>% 
  mutate(result = ifelse(baker == "Diana" & series >= 5 & episode >= 6, NA, result))

check1_df = anti_join(results_df1, bakers_df, by = c("series", "baker" ))

bakers_results = left_join(results_df1, bakers_df, by = c("series", "baker" ))

final_df =
  anti_join(bakes_df, bakers_results, by = c("series", "baker", "episode"))
  
bakes_df = 
  bakes_df %>% 
  mutate(
  baker = if_else(baker == '"Jo"', "Jo", baker))

check2_df = 
  anti_join(bakes_df, bakers_results, by = c("series", "baker", "episode"))
  
final = left_join(bakers_results, bakes_df, by = c("series", "baker", "episode"))

final_obs <- final |> nrow()

write_csv(final, "./gbb_datasets/final.csv")
```
Next, I use `str()` to check the variable types before merging and make sure variables would be merged are in the same type. `Anti_join` the bakers_df and results_df first to check missing data and find `Jo` isn't in result_df but there is someone called `Joanne`, and they are all in season2, they might be the same person, so I use mutate function to replace Joanne with Jo. Moreover, I find the result for Diana in episode 5 in series5 is `WD`, but the results for following episode and series still indicate she stayed in. So I convert the IN to NA. Then I use `left_join` to merge these two data frames. `Anti-join` the first merging data frame with bakes_df, Jo's name (8 observations) was quoted, so I remove the quote sign and then `anti_join` again, check there is 0 observation of the new data frame. Finally, merging the three dataframes which totally include `r final_obs` observations.

```{r}
win_table = final %>% 
  filter(result == "STAR BAKER" | result == "WINNER") %>% 
  filter(series %in% c("5", "6", "7", "8", "9", "10")) %>% 
  select(- c(baker_occupation, hometown, signature_bake, show_stopper))
```
I kept `series`, `episode`, `baker`, `result`, `technical` and `baker_age` in the table. And we can find the winners have very high technical rank and they usually win the star baker in several episodes in that season.But surprisingly, David, the winner in season 10 never won star baker before. 

```{r}
viewers_df = read_csv("./gbb_datasets/viewers.csv",
                      na = c("NA", "", "."), show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  head(10)

mean_s1 =
  viewers_df %>% 
  pull(series_1) %>% 
  mean(na.rm = TRUE)

mean_s5 =
  viewers_df %>% 
  pull(series_5) %>% 
  mean(na.rm = TRUE)
```
The average viewership is `r mean_s1` in Season 1 and `r mean_s5` in Season 5. 

 






